Estimating duration in the presence of misclassification

### 2024-08-01 introduction.tex
give an overview of the English pandemic

>>>>

On 31 Dec 2019, China informed the WHO (World Health Organization) of cases of pneumonia of unknown cause in Wuhan City, Hubei province~\autocite{whoNovelPneumonia2020}.
By 7 Jan 2020, the cause of the pneumonia had been identified as a novel coronavirus which would go on to be named SARS-CoV-2~\autocite{whoNovelCov2020}.
SARS-CoV-2 spread rapidly around the world: on 31 Jan 2020 the first cases were confirmed in England in two Chinese nationals~\autocite{bbcFirstUKCOVID}.
However, introductions from Spain and France in Mar 2020 were the most important sources of the epidemic in England~\autocite{duplessisEstablishment}.

The evolution of the epidemic in England was characterised by a succession of waves, with the context affected by three major types of changes: the imposition and relaxation of non-pharmaceutical interventions (NPIs), the emergence of new variants of the virus, and the roll-out of vaccination.
There were many NPIs, but the most dramatic were the three national lockdowns imposed on 23 Mar 2020, 5 Nov 2020, and 4 Jan 2021 which legally required everyone to stay at home except for limited exceptions~\autocite{ifgTimeline}.
Variants are genetic changes in the virus, the most consequential of which were the three variants of concern, named Alpha, Delta, and Omicron, that spread widely in England due to their growth advantage; prior to the emergence of Alpha, there were many variants in circulation, collectively known as ``pre-Alpha variants''~\autocite{lythgoeLineage}.
Alpha became the dominant variant in Dec 2020, Delta in May 2021, and Omicron in Dec 2021~\autocite{lythgoeLineage}.
The first vaccination in England was on 8 Dec 2020, with 68\% of adults having received two doses by Jul 2020 and 85\% by Oct 2021; initially, two doses was a full course, but later additional doses were offered, especially following the rise of the Omicron variant~\autocite{naoCovidVaccination}.
